Open source landscape ecology tools \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2020-10-11 paper/paper.Rmd
Add further results of survery

---
title: Open source landscape ecology tools
# subtitle: Do you have a subtitle? If so, write it here
# titlerunning: Open source landscape ecology tools
# authorrunning: Short form of author list if too long for running head
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.
authors: 
- name: Maximillian H.K. Hesselbarth
  address: Department of Ecosystem Modelling, University of Goettingen, Buesgenweg 4, 37077 Goettingen, Germany
  email: mhk.hesselbarth@gmail.com
- name: Jakub Nowosad
  address: Institute of Geoecology and Geoinformation, Adam Mickiewicz University, Krygowskiego 10, 61-680 Poznan, Poland
  email: nowosad.jakub@gmail.com
- name: Author 3
  address: Adress author 3
  email: E-Mail author 3
- name: ...
  address: ...
  email: ...
keywords:
- keyword 1
- keyword 2
- keyword 3
- keyword 4
- ...
abstract: |
  max. (200 words) Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
bibliography: bibliography.bib
bibstyle: spphys # bibstyle options spbasic(default), spphys, spmpsci
preamble: | 
  \usepackage{caption}
  \usepackage{ltablex}
  \usepackage{lscape}
  \usepackage{multirow}
output:
  bookdown::pdf_book:
    base_format: rticles::springer_article
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

```{r setup R, include = FALSE}
# load packages
library(dplyr)
library(forcats)
library(ggplot2)
library(googlesheets4)
library(stringr)
library(tidyr)

answers <- read_sheet("https://docs.google.com/spreadsheets/d/1mV_wZWb4desjdzUWHALIB-uFRCauSTOGSvENsnkiAhI/edit#gid=1189808333")

old_names <- names(answers)

new_names <- c("time", "position", "usage_freq", "expertise", "major_topic",
               "important_tasks", "data_model", "r_packages", "own_package",
               "own_package_location", "missing_methods", "usefulness", "comments")

names(answers) <-  new_names
```

@@

\begin{center}
  \begin{table}
    \begin{tabularx}{0.9\linewidth}{XXXX}
      \caption{Overview of commonly used \textit{R} packages for landscape ecology. Packages are sorted by their major application topic. Only packages especially for spatial data and landscape ecology are included.} \\
      \hline
      Topic & Package & Description & Source \\
      \hline
      \multirow{5}{*}{Spatial data} & \textit{raster} & Methods for raster data & \cite{Hijmans2019} \\
      & \textit{stars} & Methods for spatiotemporal arrays (raster and vector data cubes) & \cite{Pebesma2019} \\
      & \textit{sp} & Methods for mainly vector data & \cite{Pebesma2005,Bivand2013} \\
      & \textit{sf} & Methods for vector data as simple features & \cite{Pebesma2018} \\
      & \textit{landscapetools} & Utility functions for landscape analysis & \cite{Sciaini2018} \\
      \hline
      \multirow{3}{*}{Creating maps} & \textit{tmap} & xxx & xxx \\
      & \textit{mapview} & xxx & xxx \\
      & \textit{leaflet} & xxx & xxx \\
      \hline
      \multirow{2}{*}{Landscape metrics} & \textit{landscapemetrics} & Calculates landscape metrics for categorical land cover maps & \cite{Hesselbarth2019a} \\
      & \textit{belg} & Calculates the Boltzmann entropy of a landscape gradient & \cite{Nowosad2019a} \\
      \hline
      \multirow{5}{*}{}Species distribution modeling (SDM) & \textit{dismo} & xxx & xxx \\
      & \textit{sdm} & xxx & xxx \\
      & \textit{ecospat} & xxx & xxx \\
      & \textit{biomod2} & xxx & xxx \\
      & \textit{PresenceAbsence} & xxx & xxx \\
      \hline
      \multirow{3}{*}{Connectivity} & \textit{grainscape} & xxx & \cite{Chubaty2020} \\
      & \textit{gdistance} & xxx & \cite{vanEtten2017} \\
      & \textit{samc} & xxx & \cite{Marx2020} \\
      \hline
      \multirow{3}{*}{Landscape genetics} & \textit{graphs4lg} & xxx & \cite{Savary2020} \\
      & \textit{PopGenReport} & xxx & \cite{Adamack2014,Gruber2015} \\
      & \textit{HierDpart} & xxx & \cite{Qin2019} \\
      \hline
      Neutral Landscape Models (NLM) & \textit{NLMR} & xxx & \cite{Sciaini2018} \\
      \hline
    \end{tabularx}
  \end{table}
\end{center}

@@

# expertise
expertise <- dplyr::group_by(answers, expertise) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)
```

To better understand how *R* is used by the landscape ecology community, we designed a short survey and used mailing lists and social media to reach the landscape ecology community. 
In total, the survey was answered by `r n_people` people, of which the the majority were either "PhD students" (`r current_position[[1, 2]]`%), followed by "Post-Docs" (`r current_position[[2, 2]]`%) and "Professors" (`r current_position[[3, 2]]`%). Other less frequent answers were "data scientists", "government employees" or "bachelor's and master's degree students". 
Most people used R either "daily" (`r usage_freq[[1, 2]]`%) or a "few times a week" (`r usage_freq[[2, 2]]`%) and and about half of all participants described themselves as "advanced" user (`r expertise[[1, 2]]`%), while still `r expertise[[2, 2]]`% described themselves as "intermediate" users.

```{r data_model_packages, include = FALSE}
# data model
data_model <- dplyr::group_by(answers, data_model) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)

@@

# reorder levels
r_packages_df$r_package_reclass <- factor(r_packages_df$r_package_reclass, 
                                          levels = new_levels)
```
Interestingly, most people used the raster data model (`r data_model[[1, 2]]`%) in comparison to the vector data model (`r data_model[[2, 2]]`%).

```{r used_package, echo = FALSE, fig.align = "center"}
ggplot(data = r_packages_df) +
  geom_bar(aes(y = r_package_reclass, x = n_rel, fill = col), stat = "identity") + 
  geom_text(aes(y = r_package_reclass, x = n_rel, label = label, col = col), 
            nudge_x = 2.5) + 
  theme_classic() + 
  scale_x_continuous(breaks = seq(0, 45, 5), limits = c(0, 45)) + 
  scale_fill_manual(values = c("grey50", "black")) +
  scale_colour_manual(values = c("grey50", "black")) +
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "Relative value [%]", y = expression(paste(italic(R), " package")))
```

>>>>

---
title: Open source landscape ecology tools
# subtitle: Do you have a subtitle? If so, write it here
# titlerunning: Open source landscape ecology tools
# authorrunning: Short form of author list if too long for running head
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.
authors: 
- name: Maximillian H.K. Hesselbarth
  address: Department of Ecosystem Modelling, University of Goettingen, Buesgenweg 4, 37077 Goettingen, Germany
  email: mhk.hesselbarth@gmail.com
- name: Jakub Nowosad
  address: Institute of Geoecology and Geoinformation, Adam Mickiewicz University, Krygowskiego 10, 61-680 Poznan, Poland
  email: nowosad.jakub@gmail.com
- name: Author 3
  address: Adress author 3
  email: E-Mail author 3
- name: ...
  address: ...
  email: ...
keywords:
- keyword 1
- keyword 2
- keyword 3
- keyword 4
- ...
abstract: |
  max. (200 words) Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
bibliography: bibliography.bib
bibstyle: spphys # bibstyle options spbasic(default), spphys, spmpsci
preamble: | 
  \usepackage{caption}
  \usepackage{ltablex}
  \usepackage{pdflscape}
  \usepackage{multirow}
  \usepackage{placeins}
output:
  bookdown::pdf_book:
    base_format: rticles::springer_article
    number_sections: yes
    citation_package: default
editor_options: 
  chunk_output_type: console
---

```{r setup R, include = FALSE}
# load packages
library(dplyr)
library(forcats)
library(ggplot2)
library(googlesheets4)
library(patchwork)
library(stringr)
library(tibble)
library(tidyr)

# MAKE SURE TO RUN R/get_survey_data.R TO GET NEWST SURVERY RESULTS

results <- readRDS("results.rds")

answers <- results$answers

old_names <- results$old_names

new_names <- results$new_names
```

@@

  \begin{table}
    \begin{tabularx}{0.9\linewidth}{XXXX}
      \caption{Overview of commonly used \textit{R} packages for landscape ecology. Packages are sorted by their major application topic. Only packages especially for spatial data and landscape ecology are included.} \\
      \hline
      Topic & Package & Description & Source \\
      \hline
      \multirow{5}{*}{Spatial data} & \textit{raster} & Methods for raster data & \cite{Hijmans2019} \\
      & \textit{stars} & Methods for spatiotemporal arrays (raster and vector data cubes) & \cite{Pebesma2019} \\
      & \textit{sp} & Methods for mainly vector data & \cite{Pebesma2005,Bivand2013} \\
      & \textit{sf} & Methods for vector data as simple features & \cite{Pebesma2018} \\
      & \textit{landscapetools} & Utility functions for landscape analysis & \cite{Sciaini2018} \\
      \hline
      \multirow{3}{*}{Creating maps} & \textit{tmap} & xxx & xxx \\
      & \textit{mapview} & xxx & xxx \\
      & \textit{leaflet} & xxx & xxx \\
      \hline
      \multirow{2}{*}{Landscape metrics} & \textit{landscapemetrics} & Calculates landscape metrics for categorical land cover maps & \cite{Hesselbarth2019a} \\
      & \textit{belg} & Calculates the Boltzmann entropy of a landscape gradient & \cite{Nowosad2019a} \\
      \hline
      \multirow{5}{*}{}Species distribution modeling (SDM) & \textit{dismo} & xxx & xxx \\
      & \textit{sdm} & xxx & xxx \\
      & \textit{ecospat} & xxx & xxx \\
      & \textit{biomod2} & xxx & xxx \\
      & \textit{PresenceAbsence} & xxx & xxx \\
      \hline
      \multirow{3}{*}{Connectivity} & \textit{grainscape} & xxx & \cite{Chubaty2020} \\
      & \textit{gdistance} & xxx & \cite{vanEtten2017} \\
      & \textit{samc} & xxx & \cite{Marx2020} \\
      \hline
      \multirow{3}{*}{Landscape genetics} & \textit{graphs4lg} & xxx & \cite{Savary2020} \\
      & \textit{PopGenReport} & xxx & \cite{Adamack2014,Gruber2015} \\
      & \textit{HierDpart} & xxx & \cite{Qin2019} \\
      \hline
      Neutral Landscape Models (NLM) & \textit{NLMR} & xxx & \cite{Sciaini2018} \\
      \hline
    \end{tabularx}
  \end{table}

@@

# expertise
expertise <- dplyr::group_by(answers, expertise) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)

# own package
own_package <- dplyr::group_by(answers, own_package) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)
```

To better understand how *R* is used by the landscape ecology community, we designed a short survey and used mailing lists and social media to reach the landscape ecology community. 
In total, the survey was answered by `r n_people` participants, of which the the majority were either "PhD students" (`r current_position[[1, 2]]`%), followed by "Post-Docs" (`r current_position[[2, 2]]`%) and "Professors" (`r current_position[[3, 2]]`%). Other less frequent answers were "data scientists", "government employees" or "bachelor's and master's degree students". 
Most people used R either "daily" (`r usage_freq[[1, 2]]`%) or a "few times a week" (`r usage_freq[[2, 2]]`%) and and about half of all participants described themselves as "advanced" user (`r expertise[[1, 2]]`%), while still `r expertise[[2, 2]]`% described themselves as "intermediate" users. 
Related to this, about half of the participants either implemented their own package (`r own_package[[3, 2]]`%), or plan to do so in the future (`r own_package[[2, 2]]`%) and most of these packages are hosted on *GitHub* and/or *CRAN*.

```{r data-model-packages, include = FALSE}
# data model
data_model <- dplyr::group_by(answers, data_model) %>% 
  dplyr::summarise(n_rel = n() / n_people * 100) %>% 
  dplyr::mutate(n_rel = round(n_rel, digits = digits)) %>% 
  dplyr::arrange(-n_rel)

@@

# reorder levels
r_packages_df$r_package_reclass <- factor(r_packages_df$r_package_reclass, 
                                          levels = new_levels)

ggplot_packages <- ggplot(data = r_packages_df) +
  geom_bar(aes(y = r_package_reclass, x = n_rel, fill = col), stat = "identity") + 
  # geom_text(aes(y = r_package_reclass, x = n_rel, label = label, col = col), 
  #           nudge_x = 2, size = 2.5) + 
  theme_classic(base_size = 7.5) + 
  scale_x_continuous(breaks = seq(0, 35, 5), limits = c(0, 35)) + 
  scale_fill_manual(values = c("grey50", "black")) +
  scale_colour_manual(values = c("grey50", "black")) +
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "Number of package mentionings [%]", y = "List the three R packages related to landscape ecology you use the most!")
```

Interestingly, most people use the raster data model (`r data_model[[1, 2]]`%) in comparison to the vector data model (`r data_model[[2, 2]]`%). 
This was also represented in the most used *R* packages (Fig. \@ref(fig:fig-packages-tasks) A))
When asked for the three most used packages, participants of the survey listed `r length(unique(r_packages))` packages in total. The *raster* packages was listed the most, followed by the *sf* package. 
Because both the *raster* and the *sf* are designed for basic and advanced data handling and processing of raster and vector data, respectively, it is not surprising that "(pre-)processing of spatial data" was the most important workflow task for the majority of the survey participants (Fig. \@ref(fig:fig-packages-tasks) B)).

```{r tasks, include = FALSE}
# get data 
important_tasks <- unlist(str_split(answers$important_tasks, pattern = ","))

# remove empty space at beginning
important_tasks <- str_trim(important_tasks)

# make sure all lower case
important_tasks <- str_to_lower(important_tasks)

# shorten labels
important_tasks[important_tasks == "(pre-)processing of spatial data"] <- "(pre-)processing data"

important_tasks[important_tasks == "spatial point pattern analysis"] <- "point pattern analysis"

important_tasks[important_tasks == "programming individual based models"] <- "individual based modelling"

important_tasks[important_tasks == "creating neutral landscape models"] <- "neutral landscape models"

important_tasks[important_tasks == "quantifying landscape characteristics (e.g. landscape metrics)"] <- "quantifying landscape characteristics"

# add line breaks
important_tasks <- str_replace_all(important_tasks, pattern = " ", replacement = "\n")

# create data.frame
tasks_df <- tibble(tasks = important_tasks) %>%
  dplyr::group_by(tasks) %>% 
  dplyr::summarise(n = dplyr::n()) %>% 
  dplyr::mutate(n_rel = n / sum(n) * 100, 
                tasks_reclass = dplyr::case_when(n >= 5 ~ tasks, 
                                                 n < 5 ~ "others")) %>%
  dplyr::group_by(tasks_reclass) %>% 
  dplyr::summarise(n_rel = sum(n_rel)) %>% 
  dplyr::mutate(label = paste0(round(n_rel, digits = 1), "%"), 
                col = dplyr::case_when(tasks_reclass == "others" ~ 0, 
                                       tasks_reclass != "others" ~ 1), 
                col = factor(col)) %>% 
  dplyr::arrange(-n_rel)

# change ordering
new_levels <- c("others", rev(tasks_df$tasks_reclass[!tasks_df$tasks_reclass == "others"]))

# reorder levels
tasks_df$tasks_reclass <- factor(tasks_df$tasks_reclass, levels = new_levels)

ggplot_tasks <- ggplot(data = tasks_df) +
  geom_bar(aes(y = tasks_reclass, x = n_rel, fill = col), stat = "identity") + 
  # geom_text(aes(y = tasks_reclass, x = n_rel, label = label, col = col), 
  #           nudge_x = 1.25, size = 2.5) + 
  theme_classic(base_size = 7.5) + 
  scale_x_continuous(breaks = seq(0, 20, 5), limits = c(0, 20)) + 
  scale_fill_manual(values = c("grey50", "black")) +
  scale_colour_manual(values = c("grey50", "black")) +
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "Number of task mentionings [%]", y = "Which are the most important tasks to your workflow?")
```

```{r fig-packages-tasks, echo = FALSE, fig.align="center", fig.cap = "A) Most used R packages and B) most important workflow tasks as listed by the participants of the survery. 'others' includes all packages/tasks with less than five total mentions."}

ggplot_packages + ggplot_tasks + plot_layout(ncol = 2, nrow = 1) + plot_annotation(tag_levels = "A", tag_suffix = ")")

```

```{r usefulness, include = FALSE}
# create vector with classses

classes <- c("very useful", "useful", "intermediate", 
             "not useful", "not useful at all")

# create data.frame
usefulness_df <- dplyr::group_by(answers, usefulness) %>% 
  dplyr::summarise(n = dplyr::n()) %>% 
  dplyr::mutate(n_rel = n / sum(n) * 100) %>%
  tibble::add_row(usefulness = 1, n = 0, n_rel = 0) %>% 
  dplyr::arrange(-usefulness) %>% 
  dplyr::mutate(label = paste0(round(n_rel, digits = 1), "%"), 
                usefulness_class = classes,
                usefulness_class = factor(usefulness_class,
                                          levels = rev(classes))) %>% 
  dplyr::arrange(-n_rel)

good_eval <- dplyr::filter(usefulness_df, usefulness %in% c(5, 4)) %>% 
  dplyr::pull(n_rel) %>% 
  sum() %>% 
  round(digits = 2)

bad_eval <- dplyr::filter(usefulness_df, usefulness %in% c(3, 2, 1)) %>% 
  dplyr::pull(n_rel) %>% 
  sum() %>% 
  round(digits = 2)
```

When asked "`r old_names[12]`", the vast majority of participants answered with either "very useful" or "useful" (`r good_eval`%) and only only very few participants evaluated *R* as "intermediate", "not useful" or "not useful at all" (`r bad_eval`%; Fig. \@ref(fig:fig-usefulness)).

```{r fig-usefulness, echo = FALSE, fig.align="center", fig.cap = "Evaluation of the usefullnes o fR for landscape ecology by the particpants of the survey. Answers were possible on a 5-point scale."}
ggplot(data = usefulness_df) +
  geom_bar(aes(x = usefulness_class, y = n_rel), 
           stat = "identity") + 
  theme_classic(base_size = 7.5) + 
  scale_y_continuous(breaks = seq(0, 60, 5), limits = c(0, 60)) + 
  guides(fill = FALSE, colour = FALSE) +
  labs(x = "How useful do you think is R currently for landscape ecology?", 
       y = "Number of participants [%]")
```

\FloatBarrier

### 2020-10-11 paper/paper.tex
Add further results of survery

%% load any required packages here
\usepackage{caption}
\usepackage{ltablex}
\usepackage{lscape}
\usepackage{multirow}

@@

To better understand how \emph{R} is used by the landscape ecology community, we designed a short survey and used mailing lists and social media to reach the landscape ecology community.
In total, the survey was answered by 94 people, of which the the majority were either ``PhD students'' (33\%), followed by ``Post-Docs'' (26.6\%) and ``Professors'' (12.8\%). Other less frequent answers were ``data scientists'', ``government employees'' or ``bachelor's and master's degree students''.
Most people used R either ``daily'' (58.5\%) or a ``few times a week'' (34\%) and and about half of all participants described themselves as ``advanced'' user (48.9\%), while still 39.4\% described themselves as ``intermediate'' users.

Interestingly, most people used the raster data model (70.2\%) in comparison to the vector data model (29.8\%).

\begin{center}\includegraphics{paper_files/figure-latex/used_package-1} \end{center}

>>>>

%% load any required packages here
\usepackage{caption}
\usepackage{ltablex}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{placeins}

@@

To better understand how \emph{R} is used by the landscape ecology community, we designed a short survey and used mailing lists and social media to reach the landscape ecology community.
In total, the survey was answered by 94 participants, of which the the majority were either ``PhD students'' (33\%), followed by ``Post-Docs'' (26.6\%) and ``Professors'' (12.8\%). Other less frequent answers were ``data scientists'', ``government employees'' or ``bachelor's and master's degree students''.
Most people used R either ``daily'' (58.5\%) or a ``few times a week'' (34\%) and and about half of all participants described themselves as ``advanced'' user (48.9\%), while still 39.4\% described themselves as ``intermediate'' users.
Related to this, about half of the participants either implemented their own package (20.2\%), or plan to do so in the future (22.3\%) and most of these packages are hosted on \emph{GitHub} and/or \emph{CRAN}.

Interestingly, most people use the raster data model (70.2\%) in comparison to the vector data model (29.8\%).
This was also represented in the most used \emph{R} packages (Fig. \ref{fig:fig-packages-tasks} A))
When asked for the three most used packages, participants of the survey listed 75 packages in total. The \emph{raster} packages was listed the most, followed by the \emph{sf} package.
Because both the \emph{raster} and the \emph{sf} are designed for basic and advanced data handling and processing of raster and vector data, respectively, it is not surprising that ``(pre-)processing of spatial data'' was the most important workflow task for the majority of the survey participants (Fig. \ref{fig:fig-packages-tasks} B)).

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig-packages-tasks-1}

}

\caption{A) Most used R packages and B) most important workflow tasks as listed by the participants of the survery. 'others' includes all packages/tasks with less than five total mentions.}\label{fig:fig-packages-tasks}
\end{figure}

When asked ``How useful do you think is R currently for landscape ecology?'', the vast majority of participants answered with either ``very useful'' or ``useful'' (91.49\%) and only only very few participants evaluated \emph{R} as ``intermediate'', ``not useful'' or ``not useful at all'' (8.51\%; Fig. \ref{fig:fig-usefulness}).

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/fig-usefulness-1}

}

\caption{Evaluation of the usefullnes o fR for landscape ecology by the particpants of the survey. Answers were possible on a 5-point scale.}\label{fig:fig-usefulness}
\end{figure}

\FloatBarrier
